MIP-GNN: A data-driven tool for guiding combinatorial solvers

### 2020-08-19 paper/main.tex
Started implementing new model.
Cont. implementing new model.

\paragraph{Related work} Discuss related work.
\begin{itemize}
	\item Most important GNN papers, 
	\item GNNs to solve MIPs/comb. optimization problems, \cite{Li+2018b}, \cite{Vel+2020}, \cite{Gasse+2019} \cite{Gup+2020} \cite{Kha+2016}, paper by Bistra Dilkina: 
	\item Papers on SAT and CSP solving (Hsu, LeSong, Aachen, Loukas \dots) \cite{Sel+2019}
\end{itemize}

@@

\begin{algorithm}[H]\mbox{\hfill}
	\\\textbf{Input:}  Training set $\cS$, $\varepsilon > 0$, number of epochs $E$, number of rounds $T$, stepsize $\eta > 0$, scaling constant $\rho$.  \\
	\textbf{Output:} A trained model $f_{\boldsymbol{\theta}} \colon \cI \to [0,1]^n$.
	\begin{algorithmic}[1]
		\State Initialize $\vec{v}_i^{(0)}$, $\vec{c}_i^{(0)}$
		\State Initialize parameters $\boldsymbol{\theta} = (\boldsymbol{\theta}_{\text{V}}$, $\boldsymbol{\theta}_{\text{C}}$, $\boldsymbol{\theta}_{\text{X}})$
		\State Set weights $\vec{w}_j^{(0)} \leftarrow 1$, $\vec{p}_j^{(0)} \leftarrow \nicefrac{1}{m}$ for $j$ in $[m]$
		\For{$e \text{ in } [E]$}
		\State Sample $I = (\vec{c}, \vec{A}, \vec{b})$ with variable bias $\vec{\bar{b}}$ uniformly and at random from $\cS$, inducing the bipartite graph $B(I)$
		\For{$t \text{ in } [T]$}
		\State $\vec{v}_i^{(t)} \leftarrow {f}^{(t)}_{\text{V}} \hspace{-1pt} \big( \big\{ \hspace{-4.4pt} \big\{ \big( \vec{v}^{(t-1)}_i, \vec{c}^{(t-1)}_j, \vec{p}^{(t-1)}_j, \vec{A}_{ji} \big)\hspace{-1pt} \colon c_j \in N(v_i) \big\} \hspace{-4.4pt} \big\} \big)$ for all $i$ in $[n]$

		\State $\vec{x}_i^{(t)} \leftarrow {f}^{(t)}_{\text{X}} \hspace{-2pt} \left(\vec{v}_i^{(t-1)} \right)$ for all $i$ in $[n]$ 
		\Statex
		\Statex
		\State $\vec{c}_j^{\hspace{1pt}(t)} \leftarrow {f}^{(t)}_{\text{C}} \hspace{-2pt} \left( \left\{ \hspace{-4.4pt} \left\{ \left( \vec{c}^{(t-1)}_j, \vec{v}^{(t)}_i, \vec{x}^{(t)}_i, \vec{A}_{ji}, \vec{b}_j \right)\hspace{-2pt} \colon i \in N(c_j) \right\} \hspace{-4.4pt} \right\} \right)$ for all $j$ in $[m]$ 	
		\State $\vec{e}_j^{(t)} \, \leftarrow \nicefrac{1}{\rho}\,\Big(\sum_{i \in N(j)} \label{key}\vec{A}_{ji}  \vec{x}_i \Big) - \vec{b}_j$
		%\Statex
		\State $\vec{w}_j^{(t)} \leftarrow ( 1 -  \eta \vec{e}^{(t)}_j) \vec{w}_j^{(t-1)} $ 
		\State $\vec{p}_j \leftarrow \nicefrac{w_j^{(t)}}{\boldsymbol{\Gamma}(t)} $, where $\boldsymbol{\Gamma}(t) = \sum_{j \in [m]} \vec{w}^{(t)}_j$
		\State Update solution $\vec{\bar{x}} \leftarrow \vec{\bar{x}} + \vec{x}$
		\EndFor
		\State $\vec{\bar{x}} \leftarrow \nicefrac{1}{T} \sum_{t \in [T]} \vec{x}(t) $
		\State $L \leftarrow L + \ell(\vec{\bar{x}}, \vec{\bar b})$ 
		\State Gradient descent on $\nicefrac{1}{T} L$ with regard to $\boldsymbol{\theta}$
		\EndFor
		\State \Return $\boldsymbol{\theta}$

	\end{algorithmic}
	\caption{Data-driven MWU for variable bias prediction.}
	\label{alg:as2}
\end{algorithm}

@@

\begin{algorithm}[H]\mbox{\hfill}
	\\\textbf{Input:}  Training set $S$, number of epochs $E$, number of layers $T$, stepsize $\eta > 0$.  \\
	\textbf{Output:} A trained model $f_{\boldsymbol{\theta}} \colon \cI \to [0,1]^n$.
	\begin{algorithmic}[1]
		\State Initialize $\vec{v}^{(0)}$ and $\vec{c}^{(0)}$

		\State Initialize $\vec{p}_i^{(0)} \leftarrow 1/m$
		\For{$e \text{ in } [E]$}
		\State Sample $I = (\vec{c}, \vec{A}, \vec{b})$ from $U(S)$ 
		\For{$t \text{ in } [T]$}
		\State $\vec{V}_j^{(t)} \leftarrow {f}^{(t)}_{\text{V}} \hspace{-1pt} \big( \big\{ \hspace{-4.4pt} \big\{ \big( \vec{V}^{(t-1)}_j, \vec{C}^{(t-1)}_j, \vec{p}^{(t-1)}_j, \vec{A}_{ij} \big)\hspace{-1pt} \mid c_i \in N(v_j) \big\} \hspace{-4.4pt} \big\} \big)$ for all $j$ in $[n]$

		\State $\vec{X}_j^{(t)} \leftarrow {f}^{(t)}_{\text{X}} \hspace{-2pt} \big(\vec{V}_i^{(t-1)} \big)$ for all $j$ in $[n]$ 
		\Statex
		\Statex
		\State $\vec{C}_j^{\hspace{1pt}(t)} \leftarrow {f}^{(t)}_{\text{C}} \hspace{-2pt} \big( \big\{ \hspace{-4.4pt} \big\{ \big( \vec{C}^{(t-1)}_j, \vec{V}^{(t)}_i, \vec{X}^{(t)}_i, \vec{A}_{ij}, \vec{b}_j \big)\hspace{-2pt} \mid v_i \in N(c_j) \big\} \hspace{-4.4pt} \big\} \big)$ for all $i$ in $[m]$ 	
		%\State $\vec{e}_j^{(t)} \leftarrow \sum_{i \in N(j)} \label{key}\vec{A}_{ji}  \vec{x}_i - \vec{b}_j$
		%\Statex
		\State $\vec{p}_{j}^{(t)} \hspace{3pt}\leftarrow \text{softmax}(f_{\text{P}}(\vec{e}^{(t)}))_{j}$
		\EndFor
		\State $\bar{\vec{x}} \leftarrow f(\mathbin\Vert_{t \in [T]} \vec{x}^{(t)})$
		\State $L \leftarrow L + \ell(\vec{\bar{x}}, \vec{\bar b})$ 
		\State Gradient descent on $L$ with regard to $\boldsymbol{\theta}$
		\EndFor
		\State \Return $\boldsymbol{\theta}$

@@

@@

\subsection{Connnection to boosting}

joint MWU and boosting

>>>>

\paragraph{Related work} Discuss related work.
\begin{itemize}
	\item GNNs to solve MIPs/comb. optimization problems, \cite{Li+2018b}, \cite{Vel+2020}, \cite{Gasse+2019} \cite{Gup+2020} \cite{Kha+2016}, paper by Bistra Dilkina: \cite{Fer+2020} \cite{Son+2020} \cite{Wil+2019} \cite{Kha+2016b}, Paper by Lodi \cite{Ben+2018}, \cite{Lar+2018}, \cite{Bon+2018}, \cite{Ben+2019}, \cite{Fis+2019}, \cite{Gup+2020}, \cite{Zar+2020},

	\item Papers on SAT and CSP solving (Hsu, LeSong, Aachen, Loukas \dots) \cite{Sel+2019}
\end{itemize}

@@

%
%\begin{algorithm}[H]\mbox{\hfill}
%	\\\textbf{Input:}  Training set $\cS$, $\varepsilon > 0$, number of epochs $E$, number of rounds $T$, stepsize $\eta > 0$, scaling constant $\rho$.  \\
%	\textbf{Output:} A trained model $f_{\boldsymbol{\theta}} \colon \cI \to [0,1]^n$.
%	\begin{algorithmic}[1]
%		\State Initialize $\vec{v}_i^{(0)}$, $\vec{c}_i^{(0)}$
%		\State Initialize parameters $\boldsymbol{\theta} = (\boldsymbol{\theta}_{\text{V}}$, $\boldsymbol{\theta}_{\text{C}}$, $\boldsymbol{\theta}_{\text{X}})$
%		\State Set weights $\vec{w}_j^{(0)} \leftarrow 1$, $\vec{p}_j^{(0)} \leftarrow \nicefrac{1}{m}$ for $j$ in $[m]$
%		\For{$e \text{ in } [E]$}
%		\State Sample $I = (\vec{c}, \vec{A}, \vec{b})$ with variable bias $\vec{\bar{b}}$ uniformly and at random from $\cS$, inducing the bipartite graph $B(I)$
%		\For{$t \text{ in } [T]$}
%		\State $\vec{v}_i^{(t)} \leftarrow {f}^{(t)}_{\text{V}} \hspace{-1pt} \big( \big\{ \hspace{-4.4pt} \big\{ \big( \vec{v}^{(t-1)}_i, \vec{c}^{(t-1)}_j, \vec{p}^{(t-1)}_j, \vec{A}_{ji} \big)\hspace{-1pt} \colon c_j \in N(v_i) \big\} \hspace{-4.4pt} \big\} \big)$ for all $i$ in $[n]$ 		
%		
%		\State $\vec{x}_i^{(t)} \leftarrow {f}^{(t)}_{\text{X}} \hspace{-2pt} \left(\vec{v}_i^{(t-1)} \right)$ for all $i$ in $[n]$ 
%		\Statex
%		\Statex
%		\State $\vec{c}_j^{\hspace{1pt}(t)} \leftarrow {f}^{(t)}_{\text{C}} \hspace{-2pt} \left( \left\{ \hspace{-4.4pt} \left\{ \left( \vec{c}^{(t-1)}_j, \vec{v}^{(t)}_i, \vec{x}^{(t)}_i, \vec{A}_{ji}, \vec{b}_j \right)\hspace{-2pt} \colon i \in N(c_j) \right\} \hspace{-4.4pt} \right\} \right)$ for all $j$ in $[m]$ 	
%		\State $\vec{e}_j^{(t)} \, \leftarrow \nicefrac{1}{\rho}\,\Big(\sum_{i \in N(j)} \label{key}\vec{A}_{ji}  \vec{x}_i \Big) - \vec{b}_j$
%		%\Statex
%		\State $\vec{w}_j^{(t)} \leftarrow ( 1 -  \eta \vec{e}^{(t)}_j) \vec{w}_j^{(t-1)} $ 
%		\State $\vec{p}_j \leftarrow \nicefrac{w_j^{(t)}}{\boldsymbol{\Gamma}(t)} $, where $\boldsymbol{\Gamma}(t) = \sum_{j \in [m]} \vec{w}^{(t)}_j$
%		\State Update solution $\vec{\bar{x}} \leftarrow \vec{\bar{x}} + \vec{x}$
%		\EndFor
%		\State $\vec{\bar{x}} \leftarrow \nicefrac{1}{T} \sum_{t \in [T]} \vec{x}(t) $
%		\State $L \leftarrow L + \ell(\vec{\bar{x}}, \vec{\bar b})$ 
%		\State Gradient descent on $\nicefrac{1}{T} L$ with regard to $\boldsymbol{\theta}$
%		\EndFor
%		\State \Return $\boldsymbol{\theta}$
%		
%	\end{algorithmic}
%	\caption{Data-driven MWU for variable bias prediction.}
%	\label{alg:as2}
%\end{algorithm}

@@

\begin{algorithm}[H]\mbox{\hfill}
	\\\textbf{Input:}  Training set $S$, number of epochs $E$, number of layers $T$.  \\
	\textbf{Output:} A trained model $f_{\boldsymbol{\theta}} \colon \cI \to [0,1]^n$.
	\begin{algorithmic}[1]
		\State Initialize $\vec{v}^{(0)}$, $\vec{c}^{(0)}$, and $\vec{p}_i^{(0)}$
		\Statex
		\For{$e \text{ in } [E]$}
		\State Sample $I = (\vec{c}, \vec{A}, \vec{b})$ from $U(S)$ 
		\For{$t \text{ in } [T]$}
		%\State \textit{Compute variable embedding:}
		\State $\vec{V}_j^{(t)} \leftarrow f_{\textsc{Var}}^{(t)} \hspace{-1pt} \big( \big\{ \hspace{-4.4pt} \big\{ \big( \vec{V}^{(t-1)}_j, \vec{C}^{(t-1)}_j, \vec{p}^{(t-1)}_j, \vec{A}_{ij} \big)\hspace{-1pt} \mid c_i \in N(v_j) \big\} \hspace{-4.4pt} \big\} \big)$ for all $j$ in $[n]$ 		
		\State $\vec{X}_j^{(t)} \leftarrow f_\textsc{Assn}^{(t)} \hspace{-2pt} \big(\vec{V}_i^{(t-1)} \big)$ for all $j$ in $[n]$ 
		\Statex
		%\State \textit{Compute constraint embedding:}
		\State $\vec{C}_j^{\hspace{1pt}(t)} \leftarrow f_\textsc{Con}^{(t)} \hspace{-2pt} \big( \big\{ \hspace{-4.4pt} \big\{ \big( \vec{C}^{(t-1)}_j, \vec{V}^{(t)}_i, \vec{X}^{(t)}_i, \vec{A}_{ij}, \vec{b}_j \big)\hspace{-2pt} \mid v_i \in N(c_j) \big\} \hspace{-4.4pt} \big\} \big)$ for all $i$ in $[m]$ 	
		%\State $\vec{e}_j^{(t)} \leftarrow \sum_{i \in N(j)} \label{key}\vec{A}_{ji}  \vec{x}_i - \vec{b}_j$
		%\Statex
		\State $\vec{p}_{j}^{(t)} \hspace{3pt}\leftarrow \text{softmax}(f_\textsc{Prob}^{(t)}(\vec{e}^{(t)}))_{j}$
		%\begin{equation*}
		%\vec{e}_j \leftarrow \nicefrac{1}{\rho}\,\Big(\sum_{i \in N(j)} \label{key}\vec{A}_{ji}  %\vec{v}_i \Big) - \vec{b}_j   
		%\end{equation*}
		\EndFor
		\State $\bar{\vec{X}}_j \leftarrow f_\textsc{Mrg}(\mathbin\Vert_{t \in [T]} \vec{X}_j^{(t)})$
		\State $L \leftarrow L + \ell(\vec{\bar{X}}, \vec{\bar b})$ 
		\State Gradient descent  with regard to $\boldsymbol{\theta}$
		\EndFor
		\State \Return $\boldsymbol{\theta}$

@@

\begin{equation}\label{err}
\vec{e}_j \leftarrow \nicefrac{1}{\rho}\,\Big(\sum_{i \in N(j)} \vec{A}_{ji}  \vec{v}_i \Big) - \vec{b}_j   
\end{equation}

@@

\subsection{Sample complexity}

\cm{Show that our architecure offers better then generalization than simple, shallow MLP.}
Proof idea: show that GNN can be seen low "degree" (assuming low nnz) compositional functions encoded as a tree

\cm{Connection to Stefanies and Keyulus reasoning paper}
